---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ericr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->


The goal of ericr is to query the ERIC database. It adds some functions to emulate a Title Abstract Keyword (TAK) query (which ERIC doesn't have a function for).

## Installation

You can install the development version of ericr from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("sjgknight/ericr")
devtools::install_github("sjgknight/ericr")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(ericr)
## basic example code
```

## Background

ERIC is a long-running database of education research, maintained by the US Institute of Education Sciences (IES). It includes both peer-reviewed journal articles and grey literature (reports, policy documents, conference papers, etc.). ERIC nearly lost funding in 2025, and may still be at risk. 

Using ERIC API documented at:
- ERIC API:     https://eric.ed.gov/?api
- Swagger docs: https://api.ies.ed.gov/eric/swagger-ui/

For Python check: https://eric.ed.gov/pdf/Using_ERIC_API_for_Research_Topics.pdf

QUERY SYNTAX:
   The ERIC API uses a Solr-style query syntax in the `search` parameter.
   Fields are scoped with field:value syntax, e.g. title:"AI" author:"Knight"
   Boolean operators: AND OR NOT (uppercase required)
   Phrase search: "artificial intelligence"



### Package development 

Package follows guidance at: 
- https://colinfay.me/build-api-wrapper-package-r/
- And https://ropensci.org/blog/2022/06/16/publicize-api-client-yes-no/ 
(or, at least, they were very helpful and any errors are my own)

See for an alternative approach to the functions: https://httr2.r-lib.org/articles/wrapping-apis.html 


In dev build with:

`attachment::att_amend_desc()` 

`devtools::check()`

`devtools::build()`

`usethis::use_testthat()`

`usethis::use_vignette("{your-package-name}")`

`usethis::use_readme_rmd()`

`devtools::build()`

## Example use

```{r examples}

# This search is:
## phrase search (quoted) for "artificial intelligence", 
## from any source with 'policy' in it, 
## and is not peer reviewed
## returning 100 rows

results <- eric_search(
     search       = '"artificial intelligence"',
     source       = "policy",
     peerreviewed = FALSE,
     rows         = 100
   )

  # -- Basic search -----------------------------------------------------------
  # NOTE: free-text search uses AND between words, NOT quoted phrases.
  # Quoted phrases only work in field-scoped searches (title=, author=, etc.)
  results <- eric_search(
    search = "artificial AND intelligence",
    rows   = 100L
  )
  # Or scope to title where phrases are supported:
  results_title <- eric_search(
    title = '"artificial intelligence"'
  )
  dplyr::glimpse(results)

  # -- Peer-reviewed only, AI + education 
  # (you can't do this just without specifying a field, title here)
  -------------------------------------
  peer_results <- eric_search(
    title       = '"artificial intelligence" AND education',
    peerreviewed = TRUE,
    rows         = 200L,
    max_results  = 1000L
  )

  # -- Documents with law/policy identifiers ----------------------------------
  policy_docs <- eric_search(
    search         = '"artificial intelligence"',
    has_policy_law = TRUE,
    max_results    = 200L
  )
  # See what laws/policies are referenced
  policy_docs |>
    dplyr::filter(!is.na(identifierslaw)) |>
    dplyr::select(title, publicationdateyear, identifierslaw)

  # -- TAK search (title OR abstract OR keyword) ------------------------------
  tak_query <- eric_tak(
    terms = c("artificial intelligence", "generative AI"),
    extra = "peerreviewed:F"
  )
  tak_results <- eric_search(.query = tak_query, max_results = 500L)

  # -- Specific grey lit source -----------------------------------------------
  cpre_results <- eric_search(
    search = '"education policy"',
    source = '"Consortium for Policy Research in Education"'
  )

  # -- AI policy specifically -- combining approaches -------------------------
  ai_policy <- eric_search(
    .query = eric_build_query(
      search         = '"artificial intelligence"',
      source         = "policy",
      peerreviewed   = FALSE,
      has_policy_law = TRUE
    ),
    max_results = 500L,
    rows        = 200L
  )

  # -- Export -----------------------------------------------------------------
  readr::write_csv(results, "eric_results.csv")



```



